An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-03 EconGrowth.Rnw
Added a file for the Math seminar presentation.

<<cobb-douglas functions, eval=TRUE>>=
####################
# Returns an nls Cobb-Douglas model (without energy) for the country specified. 
# No energy is included in the function to be fitted.
##
cdModel <- function(countryAbbrev){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # Run the non-linear least squares fit to the data. No energy term desired in the Cobb-Douglas equation.
  # Establish guess values for alpha and lambda.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.7 #0.7 gives good results for all countries.  
  start <- list(lambda=lambdaGuess, alpha=alphaGuess)
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha)
  modelCD <- nls(formula=model, data=dataTable, start = start,
                 #Include the next 3 lines to fit with constraints.
                 #algorithm = "port",
                 #lower = list(lambda=-Inf, alpha=0),
                 #upper = list(lambda=Inf, alpha=1)
                 )
  return(modelCD)
}

@@

####################
# Returns an nls Cobb-Douglas model for the country specified
# Give an energyType ("Q", "X", or "U") if you want to include an energy term
# If you supply a value for the gamma argument, a fit with fixed gamma will be provided.
# This function dispatches to cdModel, cdeModel, or cdFixedGammaModel as appropriate.
##
cobbDouglasModel <- function(countryAbbrev, energyType, gamma){
  if (missing(energyType)){
    # Fit the Cobb-Douglas model without energy.
    return(cdModel(countryAbbrev))
  } else {
    if (missing(gamma)){
      # Fit the Cobb-Douglas model with gamma as a free parameter
      return(cdeModel(countryAbbrev, energyType))
    }
    # Fit the Cobb-Douglas model with fixed value of gamma
    return(cdeFixedGammaModel(countryAbbrev, energyType, gamma))
  }
}

@@

gammaGrid <- seq(0.0, 0.95, by=0.05)
# print("US")
# models <- cdGrid("US", "Q", gammaGrid)
# SSE <- lapply(models, function(model) summary(model)$sigma)
# data <- cbind(gammaGrid, SSE)
# plot(gammaGrid, SSE)
# print(coef(models[[2]]))
# 
print("ZA")
models <- cdGrid("ZA", "Q", gammaGrid)
SSE <- lapply(models, function(model) summary(model)$sigma)
lambda <- lapply(models, function(model) coef(model)["lambda"])
alpha <- lapply(models, function(model) coef(model)["alpha"])
data <- cbind(lambda, alpha, gammaGrid, SSE)
print(data)
plot(gammaGrid, SSE)
plot(gammaGrid, lambda)
plot(gammaGrid, alpha)

>>>>

<<cobb-douglas functions, eval=TRUE>>=
####################
# Returns an nls Cobb-Douglas model (without energy) for the country specified. 
# No energy is included in the function to be fitted.
##
cdModel <- function(countryAbbrev){
  dataTable <- loadData(countryAbbrev) #Load the data that we need.
  # Run the non-linear least squares fit to the data. No energy term desired in the Cobb-Douglas equation.
  # Establish guess values for alpha and lambda.
  lambdaGuess <- 0.0 #guessing lambda = 0 means there is no technological progress.
  alphaGuess <- 0.7 #0.7 gives good results for all countries.  
  start <- list(lambda=lambdaGuess, alpha=alphaGuess)
  # Runs a non-linear least squares fit to the data. We've replaced beta with 1-alpha for simplicity.
  model <- iGDP ~ exp(lambda*iYear) * iCapStk^alpha * iLabor^(1 - alpha)
  modelCD <- nls(formula=model, data=dataTable, start = start,
                 #Include the next 3 lines to fit with constraints.
                 #algorithm = "port",
                 #lower = list(lambda=-Inf, alpha=0),
                 #upper = list(lambda=Inf, alpha=1)
                 )
# print("str(modelCD)")
# print(str(modelCD))
# print("")
# print(modelCD$)
#   # We estimated alpha with the above code, but we have beta = 1.0 - alpha. So, get 
#   # an estimate for beta and add it to the model.
#   summaryCD <- summary(modelCD)
# print("summaryCD")
# print(summaryCD)
#   dofCD <- summaryCD$df[2] # Gives the degrees of freedom for the model.
#   tvalCD <- qt(ciHalfLevel, df = dofCD)
#   alpha <- coef(modelCD)["alpha"]
#   betaEst <- deltaMethod(modelCD, "1 - alpha") # Estimates beta and its standard error (SE).
# print("betaEst")
# print(betaEst)
#   # Calculate beta and its confidence interval and report it.
#   # Now calculate a confidence interval on beta
#   betaCICD <- with(betaEst, Estimate + c(-1.0, 1.0) * tvalCD * SE); betaCICD # Gives the confidence interval on beta.  
# print("betaCICD")
# print(betaCICD)

  return(modelCD)
}

cdModel("US")

@@

####################
# Returns an nls Cobb-Douglas model for the country specified
# Give an energyType ("Q", "X", or "U") if you want to include an energy term
# If you supply a value for the gamma argument, a fit with fixed gamma will be provided.
# This function dispatches to cdModel, cdeModel, or cdFixedGammaModel based on which arguments are specified.
##
cobbDouglasModel <- function(countryAbbrev, energyType, gamma){
  if (missing(energyType)){
    # Fit the Cobb-Douglas model without energy.
    return(cdModel(countryAbbrev))
  }
  if (!missing(gamma)){
    # Fit the Cobb-Douglas model with fixed value of gamma
    return(cdeFixedGammaModel(countryAbbrev, energyType, gamma))
  }
  # Fit the Cobb-Douglas model with gamma as a free parameter
  return(cdeModel(countryAbbrev, energyType))
}

@@

gammaGrid <- seq(0.0, 0.95, by=0.05)
# print("US")
# models <- cdGrid("US", "Q", gammaGrid)
# SSE <- lapply(models, function(model) summary(model)$sigma)
# data <- cbind(gammaGrid, SSE)
# plot(gammaGrid, SSE)
# print(coef(models[[2]]))
# 
print("ZA")
models <- cdGrid("ZA", "Q", gammaGrid)
SSE <- lapply(models, function(model) summary(model)$sigma)
lambda <- lapply(models, function(model) coef(model)["lambda"])
alpha <- lapply(models, function(model) coef(model)["alpha"])
data <- cbind(lambda, alpha, gammaGrid, SSE)
plot(gammaGrid, SSE)
plot(gammaGrid, lambda)
plot(gammaGrid, alpha)
